---
title: "Comparing Four Measures of Codon Bias"
subtitle: "Full set of genes: 1620 genes, each different at the nucleotide sequence level across 22 strains\n\nPartial set of genes: 185 genes, each different at the nucleotide sequence level and identical at the amino acid sequence level across 22 strains"
output:
  html_document: default
  pdf_document: default
date: "09/11/2019"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
library(car)
library(lme4)
library(reshape2)
```

### Analysis of Codon Bias on the Full Set of Genes
Define the response variable.
```{r response_variable}
response <- read.csv('n_PPR_1620_d.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```





#### Codon Adaptation Index (CAI) with Carbone Reference Set{.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r CAI0}
explanatory1 <- read.csv('CAI_Carbone_1620_YRC_diff_d.csv')
explanatory1 <- melt(explanatory1, id.vars = "gene", variable.name = "strain", value.name = "CAI")
dataset$CAI <- explanatory1$CAI
```

##### Summary Results of Mixed Model
``` {r CAI1}
lmm1 <- lmer(logPPR ~ CAI + (1+CAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm1)
```

##### Fixed Effects Slope Significance Results
``` {r CAI2}
lmm2 <- lmer(logPPR ~ (1+CAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm1, lmm2)
```

##### Diagnostic Plots
``` {r CAI3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm1)

scatter.smooth(fitted(lmm1), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$CAI, res, col='slateblue', span=0.1, xlab='CAI (Carbone)', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
```





#### Length-Normalized Codon Adaptation Index (nlCAI) {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r nlCAI0}
explanatory2 <- read.csv('nlCAI_1620_YRC_diff_d.csv')
explanatory2 <- melt(explanatory2, id.vars = "gene", variable.name = "strain", value.name = "nlCAI")
dataset$nlCAI <- explanatory2$nlCAI
```

##### Summary Results of Mixed Model
``` {r nlCAI1}
lmm3 <- lmer(logPPR ~ nlCAI + (1+nlCAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm3)
```

##### Fixed Effects Slope Significance Results
``` {r nlCAI2}
lmm4 <- lmer(logPPR ~ (1+nlCAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm3, lmm4)
```

##### Diagnostic Plots
``` {r nlCAI3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm3)

scatter.smooth(fitted(lmm3), res, col='salmon', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$nlCAI, res, col='salmon', span=0.1, xlab='nlCAI', ylab='Residuals')
qqnorm(res, col='salmon')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='salmon')
```





#### tRNA Adaptation Index (tAI) {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r tAI0}
explanatory3 <- read.csv('tAI_1620_YRC_diff_d.csv')
explanatory3 <- melt(explanatory3, id.vars = "gene", variable.name = "strain", value.name = "tAI")
dataset$tAI <- explanatory3$tAI
```

##### Summary Results of Model
``` {r tAI1}
lmm5 <- lmer(logPPR ~ tAI + (1+tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm5)
```

##### Fixed Effects Slope Significance Results
``` {r tAI2}
lmm6 <- lmer(logPPR ~ (1+tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm5, lmm6)
```

##### Diagnostic Plots
``` {r tAI3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm5)

scatter.smooth(fitted(lmm5), res, col='yellowgreen', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$tAI, res, col='yellowgreen', span=0.1, xlab='tAI', ylab='Residuals')
qqnorm(res, col='yellowgreen')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='yellowgreen')
```





#### Normalized tRNA Adaptation Index (ntAI) {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r ntAI0}
explanatory4 <- read.csv('ntAI_1620_YRC_diff_d.csv')
explanatory4 <- melt(explanatory4, id.vars = "gene", variable.name = "strain", value.name = "ntAI")
dataset$ntAI <- explanatory4$ntAI
```

##### Summary Results of Model
``` {r ntAI1}
lmm7 <- lmer(logPPR ~ ntAI + (1+ntAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm7)
```

##### Fixed Effects Slope Significance Results
``` {r ntAI2}
lmm8 <- lmer(logPPR ~ (1+ntAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm7, lmm8)
```

##### Diagnostic Plots
``` {r ntAI3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm7)

scatter.smooth(fitted(lmm7), res, col='peru', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$ntAI, res, col='peru', span=0.1, xlab='ntAI', ylab='Residuals')
qqnorm(res, col='peru')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='peru')
rm(list=ls())
```










### Analysis of Codon Bias on Partial Set of Genes
Define the response variable.
```{r response_variablet}
response <- read.csv('n_PPR_185_d.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```





#### Codon Adaptation Index (CAI) with Carbone Reference Set{.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r CAIt0}
explanatory1 <- read.csv('CAI_Carbone_185_YRC_diff_d.csv')
explanatory1 <- melt(explanatory1, id.vars = "gene", variable.name = "strain", value.name = "CAI")
dataset$CAI <- explanatory1$CAI
```

##### Summary Results of Mixed Model
``` {r CAIt1}
lmm1 <- lmer(logPPR ~ CAI + (1+CAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm1)
```

##### Fixed Effects Slope Significance Results
``` {r CAIt2}
lmm2 <- lmer(logPPR ~ (1+CAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm1, lmm2)
```

##### Diagnostic Plots
``` {r CAIt3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm1)

scatter.smooth(fitted(lmm1), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$CAI, res, col='slateblue', span=0.1, xlab='CAI (Carbone)', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
```





#### Length-Normalized Codon Adaptation Index (nlCAI) {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r nlCAIt0}
explanatory2 <- read.csv('nlCAI_185_YRC_diff_d.csv')
explanatory2 <- melt(explanatory2, id.vars = "gene", variable.name = "strain", value.name = "nlCAI")
dataset$nlCAI <- explanatory2$nlCAI
```

##### Summary Results of Mixed Model
``` {r nlCAIt1}
lmm3 <- lmer(logPPR ~ nlCAI + (1+nlCAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm3)
```

##### Fixed Effects Slope Significance Results
``` {r nlCAIt2}
lmm4 <- lmer(logPPR ~ (1+nlCAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm3, lmm4)
```

##### Diagnostic Plots
``` {r nlCAIt3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm3)

scatter.smooth(fitted(lmm3), res, col='salmon', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$nlCAI, res, col='salmon', span=0.1, xlab='nlCAI', ylab='Residuals')
qqnorm(res, col='salmon')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='salmon')
```





#### tRNA Adaptation Index (tAI) {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r tAIt0}
explanatory3 <- read.csv('tAI_185_YRC_diff_d.csv')
explanatory3 <- melt(explanatory3, id.vars = "gene", variable.name = "strain", value.name = "tAI")
dataset$tAI <- explanatory3$tAI
```

##### Summary Results of Model
``` {r StAIt1}
lmm5 <- lmer(logPPR ~ tAI + (1+tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm5)
```

##### Fixed Effects Slope Significance Results
``` {r tAIt2}
lmm6 <- lmer(logPPR ~ (1+tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm5, lmm6)
```

##### Diagnostic Plots
``` {r StAIt3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm5)

scatter.smooth(fitted(lmm5), res, col='yellowgreen', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$tAI, res, col='yellowgreen', span=0.1, xlab='tAI', ylab='Residuals')
qqnorm(res, col='yellowgreen')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='yellowgreen')
```





#### Normalized tRNA Adaptation Index (ntAI) {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r ntAIt0}
explanatory4 <- read.csv('ntAI_185_YRC_diff_d.csv')
explanatory4 <- melt(explanatory4, id.vars = "gene", variable.name = "strain", value.name = "ntAI")
dataset$ntAI <- explanatory4$ntAI
```

##### Summary Results of Model
``` {r ntAIt1}
lmm7 <- lmer(logPPR ~ ntAI + (0+ntAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm7)
```

##### Fixed Effects Slope Significance Results
``` {r ntAIt2}
lmm8 <- lmer(logPPR ~ (0+ntAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm7, lmm8)
```

##### Diagnostic Plots
``` {r ntAIt3, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm7)

scatter.smooth(fitted(lmm7), res, col='peru', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$ntAI, res, col='peru', span=0.1, xlab='ntAI', ylab='Residuals')
qqnorm(res, col='peru')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='peru')
```