---
title: "logPPR as a function of tAI and the deltaG of a small CDS region"
date: "03/15/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(car)
library(lme4)
library(reshape2)
```


#### logPPR ~ tAI + one-fourth CDS deltaG + interactions {.tabset .tabset-dropdown}
##### Menu
##### Define the response variable, logPPR.
```{r response_variable}
response <- read.csv("n_PPR_1322_d.csv")
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```

##### Read in 1/4 CDS deltaG Data
```{r a}
explanatory0 <- read.csv('results_onefourth_CDS_40_1322.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "onefour")
dataset$onefour = explanatory0$onefour/1000/10
```

##### Read in tAI Data
```{r d}
explanatory3 <- read.csv('tAI_1322_YRC_diff_d.csv')
explanatory3 <- melt(explanatory3, id.vars = "gene", variable.name = "strain", value.name = "tAI")
dataset$tAI = explanatory3$tAI
```

#include median protein abundance for each gene
```{r pa}
explanatory4 <- read.csv('PA_medians_1322.csv')
explanatory4 <- melt(explanatory4, id.vars = "gene", variable.name = "strain", value.name = "medianPA")
dataset$medianPA = explanatory4$medianPA/10000
head(dataset)
```

##### Summary Results of Mixed Model
``` {r F1}
lmm0 <- lmer(logPPR ~ tAI + onefour + tAI:onefour + (0+tAI|gene) + (0+onefour|gene) + (0+tAI|medianPA) + (0+onefour|medianPA) + (1|medianPA), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### 95% Confidence Intervals
``` {r F1confint}
confint(lmm0, parm="tAI", level=0.95)
confint(lmm0, parm="onefour", level=0.95)
confint(lmm0, parm="tAI:onefour", level=0.95)
```



#### logPPR ~ tAI + one-half CDS deltaG + interactions {.tabset .tabset-dropdown}
##### Menu
##### Define the response variable, logPPR.
```{r response_variable2}
response <- read.csv("n_PPR_1313_d.csv")
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```

##### Read in 1/2 CDS deltaG Data
```{r a2}
explanatory0 <- read.csv('results_onehalf_CDS_40_1313.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "onehalf")
dataset$onehalf = explanatory0$onehalf/1000/10
```


##### Read in tAI Data
```{r d2}
explanatory3 <- read.csv('tAI_1313_YRC_diff_d.csv')
explanatory3 <- melt(explanatory3, id.vars = "gene", variable.name = "strain", value.name = "tAI")
dataset$tAI = explanatory3$tAI
```

#include median protein abundance for each gene
```{r pa2}
explanatory4 <- read.csv('PA_medians_1313.csv')
explanatory4 <- melt(explanatory4, id.vars = "gene", variable.name = "strain", value.name = "medianPA")
dataset$medianPA = explanatory4$medianPA/10000
head(dataset)
```

##### Summary Results of Mixed Model
``` {r F12}
lmm0 <- lmer(logPPR ~ tAI + onehalf + tAI:onehalf + (0+tAI|gene) + (0+tAI|medianPA) + (0+onehalf|medianPA) + (1|medianPA), data=dataset, REML = FALSE, control = lmerControl(optimizer="Nelder_Mead", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```
##### 95% Confidence Intervals
``` {r F1confint2}
confint(lmm0, parm="tAI", level=0.95)
confint(lmm0, parm="onehalf", level=0.95)
confint(lmm0, parm="tAI:onehalf", level=0.95)
```

#### logPPR ~ tAI + three-fourth CDS deltaG  + interactions {.tabset .tabset-dropdown}
##### Menu
##### Define the response variable, logPPR.
```{r response_variable3}
response <- read.csv("n_PPR_1314_d.csv")
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```

##### Read in 3/4 CDS deltaG Data
```{r a3}
explanatory0 <- read.csv('results_threefourth_CDS_40_1314.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "threefour")
dataset$threefour = explanatory0$threefour/1000/10
```


##### Read in tAI Data
```{r d3}
explanatory3 <- read.csv('tAI_1314_YRC_diff_d.csv')
explanatory3 <- melt(explanatory3, id.vars = "gene", variable.name = "strain", value.name = "tAI")
dataset$tAI = explanatory3$tAI
```

#include median protein abundance for each gene
```{r pa3}
explanatory4 <- read.csv('PA_medians_1314.csv')
explanatory4 <- melt(explanatory4, id.vars = "gene", variable.name = "strain", value.name = "medianPA")
dataset$medianPA = explanatory4$medianPA/10000
head(dataset)
```

##### Summary Results of Mixed Model
``` {r F13}
lmm0 <- lmer(logPPR ~ tAI + threefour + tAI:threefour + (0+tAI|gene) + (1|gene) + (0+tAI|medianPA) + (0+threefour|medianPA) + (1|medianPA), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### 95% Confidence Intervals
``` {r F1confint3}
confint(lmm0, parm="tAI", level=0.95)
confint(lmm0, parm="threefour", level=0.95)
confint(lmm0, parm="tAI:threefour", level=0.95)
```