---
title: "logPPR as a function of mfe deltaG and mean bp probability (synonymous SNPs only)"
date: "11/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(car)
library(lme4)
library(reshape2)
```


#### logPPR ~ mfe deltaG {.tabset .tabset-dropdown}
##### Menu
##### Define the response variable, logPPR.
```{r response_variable0}
response <- read.csv("C:/Users/anastacia/Desktop/n_PPR_176_d.csv")
dataset  <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```

##### Read in deltaG Data
```{r FdelG}
explanatory1 <- read.csv('C:/Users/anastacia/Desktop/optimal_struc_mfe_176_30C.csv')
explanatory1 <- melt(explanatory1, id.vars = "gene", variable.name = "strain", value.name = "mfeG")
dataset$mfeG = explanatory1$mfeG/10000 #divide resulting slope by 10 to get units of 1/(Mcal per mol)
```

##### Summary Results of Mixed Model
``` {r F1}
lmm0 <- lmer(logPPR ~ mfeG + (1+mfeG|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### 95% Confidence Interval of Mixed Model
``` {r F1confint}
confint(lmm0, parm="mfeG", level=0.95)
```

##### Fixed Effects Slope Significance Results
``` {r F2}
lmm00 <- lmer(logPPR ~ (1+mfeG|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm00)
```


#### logPPR ~ bp prob {.tabset .tabset-dropdown}
##### Menu
##### Define the response variable, logPPR.
```{r response_variable0X}
response <- read.csv("C:/Users/anastacia/Desktop/n_PPR_176_d.csv")
dataset  <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$logPPR <- log(dataset$PPR)
```

##### Read in deltaG Data
```{r FdelGX}
explanatory1 <- read.csv('C:/Users/anastacia/Desktop/mean_gene_bp_prob_176_30C.csv')
explanatory1 <- melt(explanatory1, id.vars = "gene", variable.name = "strain", value.name = "prob")
dataset$prob = explanatory1$prob*10
```

##### Summary Results of Mixed Model
``` {r F1X}
lmm0 <- lmer(logPPR ~ prob + (0+prob|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### 95% Confidence Interval of Mixed Model
``` {r F1confintX}
confint(lmm0, parm="prob", level=0.95)
```

##### Fixed Effects Slope Significance Results
``` {r F2X}
lmm00 <- lmer(logPPR ~ (0+prob|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm00)
```