---
title: "tAI vs. sqrtPPR and deltaG vs. sqrtPPR analyses for genes ranked by low, medium, and high tAI"
subtitle: "1447 genes, each with one or more alleles across 22 strains"
output:
  html_document: default
  pdf_document: default
date: "07/19/2022"
#rerunning anastacia's models with sqrtPPR as the response instead of logPPR (and deltaG only scaled by 1/1000)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(car)
library(lme4)
library(reshape2)
```

### Analysis of Lowly Biased Genes
Define the response variable.
```{r response_variable0}
response <- read.csv('n_PPR_723_d_lowtAI.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrtPPR <- (dataset$PPR)^(1/2)
```


#### deltaG {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r deltaG_low_tAI0}
explanatory0 <- read.csv('ensemble_fe_723_30C_lowtAI.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "deltaG_low_tAI")
dataset$deltaG_low_tAI <- explanatory0$deltaG_low_tAI/1000 #don't scale resulting slope becuase already in Mcal/mol
```

##### Summary Results of Mixed Model
``` {r deltaG_low_tAI1}
lmm0 <- lmer(sqrtPPR ~ deltaG_low_tAI + (1+deltaG_low_tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="Nelder_Mead", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### Fixed Effects Slope Significance Results
``` {r deltaG_low_tAI2}
lmm1 <- lmer(sqrtPPR ~ (1+deltaG_low_tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm1)
```

##### 95% Confidence Interval of Mixed Model
``` {r deltaG_low_tAI3}
confint(lmm0, parm="deltaG_low_tAI", level=0.95)
```

##### Diagnostic Plots
``` {r deltaG_low_tAI4, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm0)

scatter.smooth(fitted(lmm0), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$deltaG_low_tAI, res, col='slateblue', span=0.1, xlab='deltaG of low tAI genes', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
rm(list = ls(all.names = TRUE))
```





### Analysis of Mediumly Biased Genes
Define the response variable.
```{r response_variable1}
response <- read.csv('n_PPR_723_d_medtAI.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrtPPR <- (dataset$PPR)^(1/2)
```


#### deltaG {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r deltaG_med_tAI0}
explanatory0 <- read.csv('ensemble_fe_723_30C_medtAI.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "deltaG_med_tAI")
dataset$deltaG_med_tAI <- explanatory0$deltaG_med_tAI/1000 #don't scale resulting slope becuase already in Mcal/mol
```

##### Summary Results of Mixed Model
``` {r deltaG_med_tAI1}
lmm0 <- lmer(sqrtPPR ~ deltaG_med_tAI + (1+deltaG_med_tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="Nelder_Mead", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### Fixed Effects Slope Significance Results
``` {r deltaG_med_tAI2}
lmm1 <- lmer(sqrtPPR ~ (1+deltaG_med_tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm1)
```

##### 95% Confidence Interval of Mixed Model
``` {r deltaG_med_tAI3}
confint(lmm0, parm="deltaG_med_tAI", level=0.95)
```

##### Diagnostic Plots
``` {r deltaG_med_tAI4, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm0)

scatter.smooth(fitted(lmm0), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$deltaG_med_tAI, res, col='slateblue', span=0.1, xlab='deltaG of medium tAI genes', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
rm(list = ls(all.names = TRUE))
```





### Analysis of Highly Biased Genes
Define the response variable.
```{r response_variable2}
response <- read.csv('n_PPR_724_d_hightAI.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrtPPR <- (dataset$PPR)^(1/2)
```


#### deltaG {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r deltaG_high_tAI0}
explanatory0 <- read.csv('ensemble_fe_724_30C_hightAI.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "deltaG_high_tAI")
dataset$deltaG_high_tAI <- explanatory0$deltaG_high_tAI/1000 #don't scale resulting slope becuase already in Mcal/mol
```

##### Summary Results of Mixed Model
``` {r deltaG_high_tAI1}
lmm0 <- lmer(sqrtPPR ~ deltaG_high_tAI + (1+deltaG_high_tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="Nelder_Mead", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### Fixed Effects Slope Significance Results
``` {r deltaG_high_tAI2}
lmm1 <- lmer(sqrtPPR ~ (1+deltaG_high_tAI|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="Nelder_Mead", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm1)
```

##### 95% Confidence Interval of Mixed Model
``` {r deltaG_high_tAI3}
confint(lmm0, parm="deltaG_high_tAI", level=0.95)
```

##### Diagnostic Plots
``` {r deltaG_high_tAI4, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm0)

scatter.smooth(fitted(lmm0), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$deltaG_high_tAI, res, col='slateblue', span=0.1, xlab='deltaG of high tAI genes', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
rm(list = ls(all.names = TRUE))
```
