---
title: "tAI vs. sqrtPPR and deltaG vs. sqrtPPR analyses for genes ranked by low, medium, and high deltaG"
subtitle: "1447 genes, each with one or more alleles across 22 strains"
output:
  html_document: default
  pdf_document: default
date: "07/19/2022"
#rerunning anastacia's models with sqrtPPR as the response instead of logPPR (and deltaG only scaled by 1/1000)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(car)
library(lme4)
library(reshape2)
```

### Analysis of Lowly Folded Genes
Define the response variable.
```{r response_variable0}
response <- read.csv('n_PPR_723_d_lowG.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrt_PPR <- (dataset$PPR)^(1/2)
```

#### tAI {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r tAI_low_G0}
explanatory0 <- read.csv('tAI_723_YRC_diff_d_lowG.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "tAI_low_G")
dataset$tAI_low_G <- explanatory0$tAI_low_G
```

##### Summary Results of Mixed Model
``` {r tAI_low_G1}
lmm0 <- lmer(sqrt_PPR ~ tAI_low_G + (1+tAI_low_G|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### Fixed Effects Slope Significance Results
``` {r tAI_low_G2}
lmm1 <- lmer(sqrt_PPR ~ (1+tAI_low_G|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm1)
```

##### 95% Confidence Interval of Mixed Model
``` {r tAI_low_G3}
confint(lmm0, parm="tAI_low_G", level=0.95)
```

##### Diagnostic Plots
``` {r tAI_low_G4, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm0)

scatter.smooth(fitted(lmm0), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$tAI_low_G, res, col='slateblue', span=0.1, xlab='tAI of low deltaG genes', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
```




### Analysis of Mediumly Folded Genes
Define the response variable.
```{r response_variable1}
response <- read.csv('n_PPR_723_d_medG.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrt_PPR <- (dataset$PPR)^(1/2)
```

#### tAI {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r tAI_med_G0}
explanatory0 <- read.csv('tAI_723_YRC_diff_d_medG.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "tAI_med_G")
dataset$tAI_med_G <- explanatory0$tAI_med_G
```

##### Summary Results of Mixed Model
``` {r tAI_med_G1}
lmm0 <- lmer(sqrt_PPR ~ tAI_med_G + (1+tAI_med_G|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### Fixed Effects Slope Significance Results
``` {r tAI_med_G2}
lmm1 <- lmer(sqrt_PPR ~ (1+tAI_med_G|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm1)
```

##### 95% Confidence Interval of Mixed Model
``` {r tAI_med_G3}
confint(lmm0, parm="tAI_med_G", level=0.95)
```

##### Diagnostic Plots
``` {r tAI_med_G4, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm0)

scatter.smooth(fitted(lmm0), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$tAI_med_G, res, col='slateblue', span=0.1, xlab='tAI of medium deltaG genes', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
```





### Analysis of Highly Folded Genes
Define the response variable.
```{r response_variable2}
response <- read.csv('n_PPR_724_d_highG.csv')
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrt_PPR <- (dataset$PPR)^(1/2)
```

#### tAI {.tabset .tabset-dropdown}
##### Menu
##### Read in Data
```{r tAI_high_G0}
explanatory0 <- read.csv('tAI_724_YRC_diff_d_highG.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "tAI_high_G")
dataset$tAI_high_G <- explanatory0$tAI_high_G
```

##### Summary Results of Mixed Model
``` {r tAI_high_G1}
lmm0 <- lmer(sqrt_PPR ~ tAI_high_G + (1+tAI_high_G|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### Fixed Effects Slope Significance Results
``` {r tAI_high_G2}
lmm1 <- lmer(sqrt_PPR ~ (1+tAI_high_G|gene), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
anova(lmm0, lmm1)
```

##### 95% Confidence Interval of Mixed Model
``` {r tAI_high_G3}
confint(lmm0, parm="tAI_high_G", level=0.95)
```

##### Diagnostic Plots
``` {r tAI_high_G4, fig.align='center', fig.height=10, fig.width=10}
par(mfrow=c(2,2))
res = residuals(lmm0)

scatter.smooth(fitted(lmm0), res, col='slateblue', span=0.1, xlab='Fitted Values', ylab='Residuals')
scatter.smooth(dataset$tAI_high_G, res, col='slateblue', span=0.1, xlab='tAI of high deltaG genes', ylab='Residuals')
qqnorm(res, col='slateblue')
qqline(res)
hist(res, main='', xlab='Residuals', ylab='Frequency', col='slateblue')
```
