---
title: "small_sections_deltaG_tAI_interaction  with sqrtPPR as response variable"
date: "3/15/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
library(car)
library(lme4)
library(reshape2)
```


#### sqrtPPR ~ sectional deltaG's + taI + all pairwise interactions{.tabset .tabset-dropdown}
##### Menu
##### Define the response variable, sqrtPPR.
```{r response_variable}
response <- read.csv("n_PPR_774_d.csv")
dataset <- melt(response, id.vars = "gene", variable.name = "strain", value.name = "PPR")
dataset$sqrtPPR <- sqrt(dataset$PPR)
```

##### Read in cap deltaG Data
```{r a}
explanatory0 <- read.csv('cap_struc_frac_30C_G_774.csv')
explanatory0 <- melt(explanatory0, id.vars = "gene", variable.name = "strain", value.name = "cap")
dataset$cap = explanatory0$cap/1000
```

##### Read in upsteamAUG deltaG Data
```{r b}
explanatory1 <- read.csv('upA_struc_frac_30C_G_774.csv')
explanatory1 <- melt(explanatory1, id.vars = "gene", variable.name = "strain", value.name = "upA")
dataset$upA = explanatory1$upA/1000
```

##### Read in downstreamAUG deltaG Data
```{r c}
explanatory2 <- read.csv('dwA_struc_frac_30C_G_774.csv')
explanatory2 <- melt(explanatory2, id.vars = "gene", variable.name = "strain", value.name = "dwA")
dataset$dwA = explanatory2$dwA/1000
```

##### Read in downstreamSTOP deltaG Data
```{r d}
explanatory3 <- read.csv('dwS_struc_frac_30C_G_774.csv')
explanatory3 <- melt(explanatory3, id.vars = "gene", variable.name = "strain", value.name = "dwS")
dataset$dwS = explanatory3$dwS/1000
```

##### Read in tAI Data
```{r e}
explanatory4 <- read.csv('tAI_774_YRC_diff_d.csv')
explanatory4 <- melt(explanatory4, id.vars = "gene", variable.name = "strain", value.name = "tAI")
dataset$tAI = explanatory4$tAI
```

#include median protein abundance for each gene
```{r pa}
explanatory5 <- read.csv('PA_medians_774.csv')
explanatory5 <- melt(explanatory5, id.vars = "gene", variable.name = "strain", value.name = "medianPA")
dataset$medianPA = explanatory5$medianPA/10000
head(dataset)
```

##### Summary Results of Mixed Model
``` {r F1}
lmm0 <- lmer(sqrtPPR ~ tAI + cap + upA + dwA + dwS + tAI:cap + tAI:upA + tAI:dwA + tAI:dwS + (0+tAI|gene) + (0+cap|gene) + (0+dwA|gene) + (0+tAI|medianPA) + (0+upA|medianPA) + (0+dwS|medianPA) + (1|medianPA), data=dataset, REML = FALSE, control = lmerControl(optimizer="bobyqa", optCtrl=list(maxfun=1e5)))
summary(lmm0)
```

##### 95% Confidence Intervals
``` {r F1confint}
confint(lmm0, parm="cap", level=0.95)
confint(lmm0, parm="upA", level=0.95)
confint(lmm0, parm="dwA", level=0.95)
confint(lmm0, parm="dwS", level=0.95)
confint(lmm0, parm="tAI", level=0.95)
confint(lmm0, parm="tAI:cap", level=0.95)
confint(lmm0, parm="tAI:upA", level=0.95)
confint(lmm0, parm="tAI:dwA", level=0.95)
confint(lmm0, parm="tAI:dwS", level=0.95)
```